DATA DRIVEN MODELS FOR APOGEE M DWARFS

### 2018-03-30 paper/draft1/draft1 (copy).tex
new figures

 (1) model the spectra using atmospheric models (BTSettl, \citealt{Allard:2011}; PHOENIX, \citealt{Husser:2013}, MARCS \citealt{Gustafsson:2008}, ATLAS \citealt{Castelli:2004}, etc.) find some places in literature where people have tried these on M dwarfs (\color{red}{Need citations... }\color{black} look at Burgasser, Blake...) - however missing opacities, and complexities due to molecules/clouds/chemistry make this difficult (\color{red}{Need citations... }\color{black} \citealt{Allard:2013});

@@

The APOGEE survey \citep{Majewski:2015} of the Sloan Digital Sky Survey IV (SDSS-IV) is a high resolution (R$\sim$22,500), H-band (1.5-1.7$\mu$m) survey which has observed over 260,000 stellar spectra since its fourteenth data release (DR14; \citealt{Abolfathi:2017}). Fundamental parameters for each of these stars are estimated by the APOGEE Stellar Parameter and Chemical Abundances Pipeline (ASPCAP; \citealt{Perez:2016}), which employs a $\chi^2$ fitting procedure using the FERRE code to fit ATLAS9 radiative transfer models \citep{Castelli:2004} to determine atmospheric parameters, 15 chemical abundances and micro-turbulence parameters (T$_{\rm eff}$, log g, [Fe/H], [α/M], [C/M], [N/M]; \citealt{Meszaros:2012}).

@@

The spectral type sample consists of 51 sources, spanning M0-M9 (Figure \ref{fig:train_dist}) cross-matched from the \citealt{West:2011} catalog of 78,841 M dwarfs from SDSS. For each source in the catalog, spectral types were determined both through an automated routine using The Hammer \citep{Covey:2007} and by visual inspection.

The physical parameter sample consists of 30 sources distributed over $3072 < T_{\rm eff} < 4131K$, and -0.48 $<$ [Fe/H] $<$ 0.49 (Figure \ref{fig:train_dist}) drawn from \citealt{Mann:2015}.
\color{red}{why logg not included as atmospheric parameter?}\color{black}

@@

\subsection{Data Processing}
Describe ASPCAP spectra files and Ness continuum normalization procedure.

>>>>

 (1) model the spectra using atmospheric models (BTSettl, \citealt{Allard:2011}; PHOENIX, \citealt{Husser:2013}, MARCS \citealt{Gustafsson:2008}, ATLAS \citealt{Castelli:2004}, etc.) find some places in literature where people have tried these on M dwarfs (\citealt{Rajpurohit:2014}, \citealt{Rajpurohit:2018}) - however missing opacities, and complexities due to molecules/clouds/chemistry make this difficult (\citealt{Allard:2013});

@@

The APOGEE survey \citep{Majewski:2015} of the Sloan Digital Sky Survey IV (SDSS-IV) is a high resolution (R$\sim$22,500), H-band (1.5-1.7$\mu$m) survey which has observed over 260,000 stellar spectra since its fourteenth data release (DR14; \citealt{Abolfathi:2017}). Fundamental parameters for each of these stars are estimated by the APOGEE Stellar Parameter and Chemical Abundances Pipeline (ASPCAP; \citealt{Perez:2016}), which employs a $\chi^2$ fitting procedure using the FERRE code to fit ATLAS9 radiative transfer models \citep{Castelli:2004} to determine atmospheric parameters, 15 chemical abundances and micro-turbulence parameters (T$_{\rm eff}$, log g, [Fe/H], [$\alpha$/M], [C/M], [N/M]; \citealt{Meszaros:2012}).

@@

The spectral type sample consists of 51 sources, spanning M0-M9 (Figure \ref{fig:train_dist}) cross-matched from the \citealt{West:2011} (hereafter W11) catalog of 78,841 M dwarfs from SDSS. For each source in the catalog, spectral types were determined both through an automated routine using The Hammer \citep{Covey:2007} and by visual inspection to a reported accuracy of $\pm$1 type.

The physical parameter sample consists of 30 sources with reference labels distributed over $3072 < T_{\rm eff} < 4131K$, and -0.48 $<$ [Fe/H] $<$ 0.49 (Figure \ref{fig:train_dist}) drawn from \citealt{Mann:2015} (hereafter M15). The M15 catalog in total contains 183 sources selected from CONCH-SHELL \citep{Gaidos:2013} and \citealt{Lepine:2011} \color{red}{these have optical specs?}\color{black} with infrared spectra taken in follow up from the SpeX Spectograph. Effective temperatures are measured by comparison of BT-SETTL atmospheric models \citep{Allard:2011} to optical spectra. A subsample of 29 sources with measured angular diameters and bolometric fluxes

Spectral types reported by M15 for the sources using optical molecular band indices (by the method of \citealt{Lepine:2013}) and ranged K7.3--M4.9 for the subsample we used, although these labels were not included in the model training.

\color{red}{why logg not included as atmospheric parameter?}\color{black}

@@

\subsection{Data Processing}

APOGEE data products come in several types of files: ap1D, apVisit, apStar, and aspcapStar.

Describe ASPCAP spectra files and Ness continuum normalization procedure.
